---
title: "Getting started with TSP"
author: "Michael Hahsler and Kurt Hornik"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Getting started with TSP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 5
)
library(TSP)
```

The traveling salesperson problem asks for the shortest round trip that visits
each city exactly once. The **TSP** package provides classes for representing
traveling salesperson problems, several construction and improvement
heuristics, and interfaces to the external Concorde solver.

This vignette covers the typical workflow:

1. Create a problem from coordinates or a distance matrix.
2. Find a tour with `solve_TSP()`.
3. Inspect, compare, and visualize solutions.

## Installation

Install the released version from CRAN:

```{r install, eval=FALSE}
install.packages("TSP")
```

Then load the package:

```{r load-package}
library(TSP)
```

## Create a problem from coordinates

For locations described by coordinates, create a Euclidean traveling
salesperson problem with `ETSP()`. Row names are used as city labels.

```{r create-etsp}
cities <- data.frame(
  x = c(0.15, 0.38, 0.62, 0.82, 0.78, 0.55, 0.28, 0.08),
  y = c(0.20, 0.08, 0.18, 0.42, 0.78, 0.92, 0.80, 0.52),
  row.names = LETTERS[1:8]
)

problem <- ETSP(cities)
problem
n_of_cities(problem)
labels(problem)
```

An `ETSP` stores the coordinates directly. Most built-in heuristics convert it
to a distance-based `TSP` internally. For very large coordinate sets this
distance matrix can require substantial memory.

## Solve the problem

`solve_TSP()` returns a `TOUR`, an integer permutation giving the order in
which the cities are visited. The default combines arbitrary insertion with
2-opt improvement.

```{r solve}
tour <- solve_TSP(problem, seed = 123)
tour
as.integer(tour)
labels(tour)
tour_length(tour)
```

The stored tour length includes the final edge from the last city back to the
first. A TSP solution is therefore a cycle, not an open path.

Plot the coordinates and the resulting tour with `plot()`:

```{r plot-tour}
plot(problem, tour, main = "A tour through eight cities", tour_col = "steelblue")
```

## Create a problem from distances

If pairwise distances or costs are already available, pass a `dist` object or
a symmetric matrix to `TSP()`.

```{r create-tsp}
distance_problem <- TSP(dist(cities))
distance_problem

distance_tour <- solve_TSP(distance_problem, method = "nn", start = 1)
distance_tour
tour_length(distance_tour)
```

`TSP()` does not require distances to be Euclidean; they can represent any
symmetric dissimilarity or cost. Missing connections can be represented by
`Inf`, while `NA` values are not allowed.

## Choose and compare heuristics

Construction heuristics build a tour from scratch. Common choices include
nearest neighbor (`"nn"`) and nearest, cheapest, farthest, or arbitrary
insertion. Improvement heuristics such as `"two_opt"` start from an existing
tour and try to shorten it.

```{r compare-methods}
methods <- c(
  "nearest neighbor" = "nn",
  "nearest insertion" = "nearest_insertion",
  "cheapest insertion" = "cheapest_insertion",
  "repetitive nearest neighbor" = "repetitive_nn"
)

tours <- lapply(methods, function(method) {
  solve_TSP(problem, method = method, seed = 123)
})

sort(vapply(tours, tour_length, numeric(1)))
```

Add `two_opt = TRUE` to refine a constructed tour, or explicitly improve an
existing tour:

```{r improve-tour}
initial_tour <- solve_TSP(problem, method = "nn", start = 1)
improved_tour <- solve_TSP(
  problem,
  method = "two_opt",
  tour = initial_tour
)

c(
  initial = tour_length(initial_tour),
  improved = tour_length(improved_tour)
)
```

Randomized methods support `seed` for reproducibility and `rep` for repeated
starts. For example, the following returns the shortest of 20 random tours,
each followed by 2-opt improvement:

```{r repeated-search}
repeated_tour <- solve_TSP(
  problem,
  method = "random",
  rep = 20,
  two_opt = TRUE,
  seed = 123
)
tour_length(repeated_tour)
```

These methods are heuristics: a short tour is not necessarily an optimal tour.
The `"concorde"` method can find an exact solution, and `"linkern"` provides
the Chained Lin-Kernighan heuristic, but both require a separately installed
Concorde executable. See `?Concorde` for setup instructions.

## Asymmetric problems

When the cost from city *i* to city *j* differs from the reverse direction,
create an asymmetric problem with `ATSP()` from a square cost matrix.

```{r create-atsp}
cost <- matrix(
  c(
    0, 4, 1, 3,
    2, 0, 5, 1,
    3, 2, 0, 4,
    6, 3, 2, 0
  ),
  nrow = 4,
  byrow = TRUE,
  dimnames = list(LETTERS[1:4], LETTERS[1:4])
)

asymmetric_problem <- ATSP(cost)
asymmetric_tour <- solve_TSP(asymmetric_problem, method = "nn", start = 1)
asymmetric_tour
tour_length(asymmetric_tour)
```

Most built-in heuristics work directly with both `TSP` and `ATSP` objects.
Solvers that only accept symmetric problems can use the package's automatic
ATSP-to-TSP reformulation; see `?reformulate_ATSP_as_TSP`.

## Where to go next

The most useful help pages are:

- `?solve_TSP` for all solvers and their control parameters;
- `?TSP`, `?ETSP`, and `?ATSP` for problem representations;
- `?TOUR` and `?tour_length` for working with solutions;
- `?read_TSPLIB` and `?write_TSPLIB` for TSPLIB files; and
- `?insert_dummy` and `?cut_tour` for turning a cycle into one or more paths.
